### 2024-01-23 outputs/paper/paper.qmd
saving what I have done so far which mostly entails my workspace set-up and introduction.

---
title: "Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
author: "Bella MacLean"
thanks: "Code and data from this analysis are available at: https://github.com/Bellamaclean7/UofT_Neighbourhood_Assaults/tree/main"
date: "today"
date-format: "long" 
abstract: "Using data from OpenDataToronto, this paper examines the trends in assault rates from 2014 to 2023 across 13 of Toronto’s 158 neighbourhoods recognized by the City of Toronto. These 13 neighbourhoods encompass and surround the University of Toronto. This papers seeks to determine a correlation between the university's presence and assault rates in its neighbouring areas, providing insights that may inform campus safety strategies and urban planning."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up ####

## Load packages
library(tidyverse)
library(knitr)
library(ggplot2)

#read in cleaned data #
uoft_neighbourhoods <- read.csv(here::here("outputs/data/cleaned_data.csv"))
show_col_types = FALSE

#read in cleaned data #
surrounding_neighbourhoods <- read.csv(here::here("outputs/data/analysis_data.csv"))
show_col_types = FALSE
```

# Introduction

Toronto is a city known for its vibrant culture and diverse neighborhoods and is home to one of Canada's most prestigious educational institutions – the University of Toronto. The University of Toronto encompasses and shapes the nearby areas of Bay-Cloverhill, The Annex, University district, and Kensington-Chinatown. These neighbourhoods are home to both an abundance of UofT academic and student buildings and a large student population. This paper examines assault rates from 2014 to 2023 in 13 of Toronto's 158 recignized neighbourhoods, using data from OpenDataToronto [@rOpenDataToronto], to understand the safety landscape of the University of Toronto in relation to assault rates over a ten-year period.

In February of 2020, the University of Toronto released data from a two-year-old province-wide, online student survey on sexual harassment and sexual assault, saying it will paint a more detailed picture of the issues and help improve supports on campus [@citeUofTSexualViolenceData]. Originally collected as part of the 2018 Student Voices on Sexual Violence Survey, the data was then released by the Ontario government to all universities and colleges province wide. Over 117,000 students from various institutions, including over 20,000 from U of T, participated in the survey. The overview of the survey released indicated that 63 per cent of students province-wide, and 59 per cent of students at U of T, reported at least one incident of sexual harassment since the beginning of the 2017-2018 academic year [@citeUofTSexualViolenceData].

Although the Univerdity of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and voilence, there is very little data and literature that exists on the matter. This paper aids to contribute to the investigation of violence in and around the Universities district.

The definition of assault in Canada under Sections 265 and 266 of the Criminal Code is any physical contact without consent or any action that causes someone to believe that such contact is imminent. This can range from an actual application of force to threatening gestures that make someone fear they might be harmed. This applies to all forms of assault, including sexual assault, sexual assault with a weapon, threats to a third party or causing bodily harm and aggravated sexual assault [@citeCriminalCodeSection265].

To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion.  In the Data section, I discuss the nature of the spreadsheet obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data. The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight. Lastly, the Conclusion section summarizes the main findings from this paper.

\newpage

# References

### 2024-01-23 scripts/02.1-data_cleaning.R
Commiting changes to neighbourhood names in the filter for cleaning the neighbourhoods in the dataset. Made cahnges in 02.1-data_cleaning.R and saved to outputs/data/analysis_data.cvs

@@ -25,7 +25,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("University", "Dovercourt Village", "Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-23 scripts/02.1-data_cleaning.R
Small correction to the filtered neighbourhoods.

@@ -25,7 +25,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-23 scripts/02.1-data_cleaning.R
Correction to the filtered neighbourhoods in the cleaned data set. Saved to outputs/data/analysis_data.csv

@@ -25,7 +25,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-23 scripts/02.1-data_cleaning.R
Renamed all the Assault Rate variables to make them look cleaner and more readable

@@ -21,7 +21,17 @@ assault_rate_data <- raw_data |>
          ASSAULT_RATE_2017, ASSAULT_RATE_2018, ASSAULT_RATE_2019, 
          ASSAULT_RATE_2020, ASSAULT_RATE_2021, ASSAULT_RATE_2022, 
          ASSAULT_RATE_2023) |>
-  rename(Neighbourhood = AREA_NAME)
+  rename(Neighbourhood = AREA_NAME,
+         `Assault Rate 2014` = ASSAULT_RATE_2014,
+         `Assault Rate 2015` = ASSAULT_RATE_2015,
+         `Assault Rate 2016` = ASSAULT_RATE_2016,
+         `Assault Rate 2017` = ASSAULT_RATE_2017,
+         `Assault Rate 2018` = ASSAULT_RATE_2018,
+         `Assault Rate 2019` = ASSAULT_RATE_2019,
+         `Assault Rate 2020` = ASSAULT_RATE_2020,
+         `Assault Rate 2021` = ASSAULT_RATE_2021,
+         `Assault Rate 2022` = ASSAULT_RATE_2022,
+         `Assault Rate 2023` = ASSAULT_RATE_2023)
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>

### 2024-01-23 scripts/02-data_cleaning.R
Renamed all the Assault Rate variables to make then look cleaner and more readable.

@@ -21,7 +21,17 @@ assault_rate_data <- raw_data |>
          ASSAULT_RATE_2017, ASSAULT_RATE_2018, ASSAULT_RATE_2019, 
          ASSAULT_RATE_2020, ASSAULT_RATE_2021, ASSAULT_RATE_2022, 
          ASSAULT_RATE_2023) |>
-  rename(Neighbourhood = AREA_NAME)
+  rename(Neighbourhood = AREA_NAME,
+         `Assault Rate 2014` = ASSAULT_RATE_2014,
+         `Assault Rate 2015` = ASSAULT_RATE_2015,
+         `Assault Rate 2016` = ASSAULT_RATE_2016,
+         `Assault Rate 2017` = ASSAULT_RATE_2017,
+         `Assault Rate 2018` = ASSAULT_RATE_2018,
+         `Assault Rate 2019` = ASSAULT_RATE_2019,
+         `Assault Rate 2020` = ASSAULT_RATE_2020,
+         `Assault Rate 2021` = ASSAULT_RATE_2021,
+         `Assault Rate 2022` = ASSAULT_RATE_2022,
+         `Assault Rate 2023` = ASSAULT_RATE_2023)
 
 # Filter rows for specific neighborhoods
 uoft_neighbourhoods <- assault_rate_data |>

### 2024-01-23 outputs/paper/paper.qmd
Saving all changes to paper.qmd that I have made thus far.

@@
 library(ggplot2)
+library(readr)
 
@@
 
-# Introduction
+## Introduction
 
@@
 
+## Data
+The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
+The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates].
+Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
+
+###Source and Data Collection
+This dataset includes the Crime Data by Neighbourhood.
+Counts are available for Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide and Shooting & Firearm Discharges.
+Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics.
+Comparing to crime count, crime rate provides a fairer comparison of the crime over time by taking into account the change in population in the region.
+The data was last refreshed on Jan 11, 2024 and is refreshed anually.
+I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year.
+I then refined the nieghbourhoods to include only those that encompass the University of Toronto and then its surrounding areas.
+The city of Toronto recognizes 158 unique Neighbourhoods and of the 158, 13 have been included in the cleaned data.
+
+###University of Toronto Neighbourhoods
+The 'Neighbourhood Crime Rates' dataset was filtered to focus solely on assault rates.
+This entailed discarding unrelated crime statistics and zeroing in on annual assault rate data per neighbourhood.
+The next crucial step was to narrow down the neighbourhoods to the four aforementioned areas central to the University of Toronto.
+These neighbourhoods are 4 of the 158 legitamate neighbourhoods recognized by the city of Toronto and can be seen in figure (PICTURE REF> HERE).
+This selective focus was vital to the study's goal of examining crime trends in university-adjacent areas (See Table \@ref(tbl-cleaned_uoft_neighbourhood_data)).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-cleaned_uoft_neighbourhood_data
+#| tbl-cap: Sample of Cleaned Data on Assault Rates in University of Toronto Neighbourhoods
+
+# Displaying the first 5 rows of the dataset
+head(cleaned_data, 5) |>
+  kable(
+    col.names = c("Neighbourhood", "Assault Rate 2014", "Assault Rate 2015", "Assault Rate 2016", "Assault Rate 2017", "Assault Rate 2018", "Assault Rate 2019", "Assault Rate 2020", "Assault Rate 2021", "Assault Rate 2022", "Assault Rate 2023"),
+    booktabs = TRUE
+  )
+```
+
+###University of Toronto Surrounding Areas
+The scope was then expanded to encompass neighbourhoods surrounding the primary four that make up the University of Toronto.
+This was done to understand the peripheral impact and the broader context of assault rates around the university.
+This secondary cleaning phase involved identifying and including adjacent neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See Table \@ref(tbl-cleaned_surrounding_neighbourhood_data)).
+These neighbourhoods make up 10 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in figure (PICTURE REF> HERE).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-cleaned_surrounding_neighbourhood_data
+#| tbl-cap: Sample of Cleaned Data on Assault Rates in Surrounding Neighbourhoods
+
+# Displaying the first 5 rows of the dataset
+head(surrounding_neighbourhoods, 5) |>
+  kable(
+    col.names = c("Neighbourhood", "Assault Rate 2014", "Assault Rate 2015", "Assault Rate 2016", "Assault Rate 2017", "Assault Rate 2018", "Assault Rate 2019", "Assault Rate 2020", "Assault Rate 2021", "Assault Rate 2022", "Assault Rate 2023"),
+    booktabs = TRUE
+  )
+```
+
+##Results
+
 # References

### 2024-01-23 outputs/paper/paper.qmd
commiting all my work thus far in paper.qmd.

@@
 library(readr)
+library(dplyr)
+library(tidyr)
 
@@
 ##Results
+###Assault Rates per University of Toronto Neighbourhood 
+Over the past decade University of Toronto's encompassing neighbourhoods have exhibited unique trends in assault rates.
+"Kensington-Chinatown" shows a erratic pattern, peaking around 2018 followed by a sharp decline in 2019 and a subsequent rise again.
+"The Annex" demonstrates a generally consitent, however, slight upward trend.
+Both "Bay-Cloverhill" and "University" show quite inconsistent patterns with several fluctuations over the ten year time span.
+What is most notable is Universities steady increase from 2020 on.
+(\@ref(fig-assaults_per_year_uoft)) shows the assault rates per year for each University of Toronto neighbourhood.
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-assaults_per_year_uoft
+#| fig-cap: Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)
+
+long_data <- cleaned_data |>
+  gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
+  mutate(Year = as.numeric(str_replace(Year, "ASSAULT_RATE_", "")))
+
+# Plot the graph
+ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(
+    title = "Assault Rates in University of Toronto Neighbourhoods (2014-2023)",
+    x = "Year",
+    y = "Assault Rate",
+    color = "Neighbourhood"
+  ) +
+  ylim(500, 2500) # Adjust these values based on your data's range
+
+```
+
+###Assault Rates per Neighbourhood in the Surrounding Areas
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-assaults_per_year_uoft
+#| fig-cap: Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)
+
+long_data <- surrounding_neighbourhoods %>%
+  gather(key = "Year", value = "Assault_Rate", -Neighbourhood) %>%
+  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")))
+
+# Check for NA values in Year
+sum(is.na(long_data$Year))
+
+# Plot the graph
+ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(
+    title = "Assault Rates in Surrounding University of Toronto Neighbourhoods (2014-2023)",
+    x = "Year",
+    y = "Assault Rate",
+    color = "Neighbourhood"
+  ) +
+  scale_x_continuous(breaks = 2014:2023) # Make sure the breaks include all years in your data
+
+```
 

### 2024-01-23 outputs/paper/paper.qmd
Made several comment updates.

@@
 #| echo: false
-#| label: fig-assaults_per_year_uoft
-#| fig-cap: Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)
+#| label: fig-assaults_per_year_surounding_neighbourhoods
+#| fig-cap: Assault Rates in Surrounding Neighbourhoods
 
@@
 
-# Check for NA values in Year
+# Checking for NA values in Year because when I was building my graph orignally the output was a graph with nothing plotted on it even though everything else looked good the Year column is producing NA values
 sum(is.na(long_data$Year))
@@
   labs(
-    title = "Assault Rates in Surrounding University of Toronto Neighbourhoods (2014-2023)",
+    title = "Assault Rates in University of Toronto's Surrounding Neighbourhoods (2014-2023)",
     x = "Year",
@@
   ) +
-  scale_x_continuous(breaks = 2014:2023) # Make sure the breaks include all years in your data
+  scale_x_continuous(breaks = seq(from = 2014, to = 2023, by = 2)) # I added this line because it specifies the break points for the x-axis, which in this context represents the years.
+This line is telling ggplot2 to place axis ticks at every integer value between 2014 and 2023.
+I had to add this because the years were not showing up there was something wrong at the x-axis was coming up blank and therefore there was no datapoints in the graph
 

### 2024-01-23 scripts/02.1-data_cleaning.R
Updated filter neighbourhoods because there was a spelling error.

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity Bellwoods", "West Queen West", "Wellington Place", "Bay-Cloverhill", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-23 scripts/02.1-data_cleaning.R
updated filtered neighbourhoods.

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Dovercourt Village", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-24 scripts/02.2_data_cleaning.R
commiting new cleaned data saved to outputs/data/discussion_data.csv

@@ -1,15 +1,47 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data to prepare for data anlysis 
+# Author: Bella MacLean
+# Date: 21 January 2024
+# Contact: bella.maclean@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: none
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
 
-#### Test data ####
+#### Clean data ####
+raw_data <- read_csv("inputs/data/raw_data.csv")
+View(raw_data)
+
+# Select the specified columns and rename AREA_NAME
+assault_rate_data <- raw_data |>
+  select(AREA_NAME, 
+         ASSAULT_RATE_2014, ASSAULT_RATE_2015, ASSAULT_RATE_2016, 
+         ASSAULT_RATE_2017, ASSAULT_RATE_2018, ASSAULT_RATE_2019, 
+         ASSAULT_RATE_2020, ASSAULT_RATE_2021, ASSAULT_RATE_2022, 
+         ASSAULT_RATE_2023) |>
+  rename(Neighbourhood = AREA_NAME,
+         `Assault Rate 2014` = ASSAULT_RATE_2014,
+         `Assault Rate 2015` = ASSAULT_RATE_2015,
+         `Assault Rate 2016` = ASSAULT_RATE_2016,
+         `Assault Rate 2017` = ASSAULT_RATE_2017,
+         `Assault Rate 2018` = ASSAULT_RATE_2018,
+         `Assault Rate 2019` = ASSAULT_RATE_2019,
+         `Assault Rate 2020` = ASSAULT_RATE_2020,
+         `Assault Rate 2021` = ASSAULT_RATE_2021,
+         `Assault Rate 2022` = ASSAULT_RATE_2022,
+         `Assault Rate 2023` = ASSAULT_RATE_2023)
+
+# Filter rows for specific neighborhoods
+all_neighbourhoods <- assault_rate_data |>
[…]

### 2024-01-24 scripts/02.2_data_cleaning.R
This was just a test.

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 all_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Kensington-Chinatown", "University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(all_neighbourhoods)

### 2024-01-24 scripts/02.2_data_cleaning.R
This is just a test (2.0).

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 all_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Kensington-Chinatown", "University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(all_neighbourhoods)

### 2024-01-24 scripts/02.1-data_cleaning.R
Made update to filtered neighbourhoods/

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(surrounding_neighbourhoods)

### 2024-01-24 scripts/02.2_data_cleaning.R
Made update to filtered neighbourhoods.

@@ -35,7 +35,7 @@ assault_rate_data <- raw_data |>
 
 # Filter rows for specific neighborhoods
 all_neighbourhoods <- assault_rate_data |>
-  filter(Neighbourhood %in% c("Kensington-Chinatown", "University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Yonge-Bay Corridor", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
+  filter(Neighbourhood %in% c("Kensington-Chinatown", "University", "Bay-Cloverhill", "Annex", "Dovercourt Village",  "Church-Wellesley", "Palmerston-Little Italy", "Trinity-Bellwoods", "West Queen West", "Wellington Place", "Wychwood", "Casa Loma", "Yonge-St.Clair"))
 
 # View the first few rows of the dataset with renamed variables
 print(all_neighbourhoods)

### 2024-01-24 outputs/paper/paper.qmd
commiting all work done in paper.qmd thus far.

--- 2024-01-23 outputs/paper/paper.qmd
@@
 library(readr)
+library(dplyr)
+library(tidyr)
+library(stringr)
 
@@
 surrounding_neighbourhoods <- read.csv(here::here("outputs/data/analysis_data.csv"))
+show_col_types = FALSE
+
+#read in cleaned data #
+all_neighbourhoods <- read.csv(here::here("outputs/data/discussion_data.csv"))
 show_col_types = FALSE
@@
 Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics.
-Comparing to crime count, crime rate provides a fairer comparison of the crime over time by taking into account the change in population in the region.
-The data was last refreshed on Jan 11, 2024 and is refreshed anually.
+Comparing to assault count, assault rate provides a fairer comparison of the crime over time by taking into account the change in population in the region.
+The data was last refreshed on Jan 11, 2024 and is refreshed annually.
 I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year.
 I then refined the nieghbourhoods to include only those that encompass the University of Toronto and then its surrounding areas.
-The city of Toronto recognizes 158 unique Neighbourhoods and of the 158, 13 have been included in the cleaned data.
+The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.
 
@@
 ##Results
+###Assault Rates per University of Toronto Neighbourhood 
+Over the past decade, University of Toronto's encompassing neighbourhoods have exhibited unique trends in assault rates.
+"Kensington-Chinatown" shows an erratic pattern, peaking around 2018 followed by a sharp decline in 2019 and a subsequent rise again.
+"The Annex" demonstrates a generally consistent, however, slight upward trend.
+Both "Bay-Cloverhill" and "University" show quite inconsistent patterns with several fluctuations over the ten year time span.
+What is most notable is Universities steady increase from 2020 on.
+The Annex also shows an increase from 2020 on, though not as steep as University.
+(\@ref(fig-assaults_per_year_uoft)) shows the assault rates per year for each University of Toronto neighbourhood.
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-assaults_per_year_uoft
+#| fig-cap: Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)
+
+long_data <- cleaned_data |>
+  gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
+  mutate(Year = as.numeric(str_replace(Year, "ASSAULT_RATE_", "")))
+
+# Plot the graph
+ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(
+    title = "Assault Rates in University of Toronto 
+      Neighbourhoods (2014-2023)",
+    x = "Year",
+    y = "Assault Rate",
+    color = "Neighbourhood"
+  ) +
+  ylim(0, 5000) # Adjust these values based on your data's range
+
+```
+
+###Assault Rates per Neighbourhood in the Surrounding Areas
+Assault rates in the neighbourhoods surrounding the University of Toronto Neighbourhoods reveals some distinct patterns and fluctuations in the data.
+Several neighbourhoods exhibit fairly stable trends with minor year-to-year changes including "Yonge-St.Clair", "Wychwood", "West Queen West, "Palmerston-Little Italy", "Case Loma" and "Trinity-Bellwoods".
+"Dovercourt Village" shows a slight upward trend over the decade.
+"Church-Wellesley" demonstrate quite a sporadic trend with several jumps and falls.
+From 2014-2020 "Wellington Place" shows a steep downward trend, rising from 2020-2022, however, then showing a steep decline again from 2022-2023.
+Several neighbourhoods exhibit quite low assault rates over the entire decade including "Yonge-St.
+Clair", "Casa Loma", "Palmerston-Little Italy", and "Wychwood".
+For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to (\@ref(fig-assaults_per_year_surounding_neighbourhoods)).
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-assaults_per_year_surounding_neighbourhoods
+#| fig-cap: Assault Rates in Surrounding Neighbourhoods
+
+long_data <- surrounding_neighbourhoods |>
+  gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
+  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")))
+
+# Checking for NA values in Year because when I was building my graph orignally the output was a graph with nothing plotted on it even though everything else looked good the Year column is producing NA values
+sum(is.na(long_data$Year))
+
+# Plot the graph
+ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(
+    title = "   Assault Rates in University of Toronto's 
+Surrounding Neighbourhoods (2014-2023)",
+    x = "Year",
+    y = "Assault Rate",
+    color = "Neighbourhood"
+  ) +
+  scale_x_continuous(breaks = seq(from = 2014, to = 2023, by = 2)) # I added this line because it specifies the break points for the x-axis, which in this context represents the years.
+This line is telling ggplot2 to place axis ticks at every integer value between 2014 and 2023.
+I had to add this because the years were not showing up there was something wrong at the x-axis was coming up blank and therefore there was no datapoints in the graph
+
+```
+#Discussion
+The data allows for a comparative analysis of assault rates across various neighborhoods in and around the University of Toronto from 2014 to 2023.
+By comparing the assault trends in the university's immediate vicinity with those of the surrounding neighborhoods, I aim to uncover underlying patterns, draw attention to areas of concern, and spark discussions around potential causative factors.
+
+The data presents a counterintuitive narrative, revealing that the University of Toronto, a Canadian fortress of education and innovation, experiences higher assault rates on average compared to its neighbouring communities.
+This finding challenges common perceptions of campus environments as insulated and safe.
+People often believe that a university is a safe bubble, away from the problems of the outside world, especially at a prestigious university focused on cutting edge learning and progress.
+However, that is not what the data portrays.
+(See\@ref(fig-assaults_per_year_per_neighbourhoods))
+
+```{r assault-rates-line-plot, fig.width=12, fig.height=8, echo=FALSE}
+#| message: false
+#| echo: false
+#| label: fig-assaults_per_year_per_neighbourhoods
+#| fig-cap: Assault Rates per year Across All Neighbourhoods (2014-2023)
+
+# Your data preparation code
+long_data <- all_neighbourhoods |>
+  gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
+  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")))
+
+# Plot the graph with facets and without a legend
+ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(
+    title = "Assault Rates in University of Toronto's Surrounding Neighbourhoods (2014-2023)",
+    x = "Year",
+    y = "Assault Rate"
+  ) +
+  scale_x_continuous(breaks = seq(from = 2014, to = 2023, by = 1)) +
+  facet_wrap(~ Neighbourhood, scales = "free_x") +
+  theme(
+    legend.position = "none", # This line removes the legend
+    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
+  )
+
+```
 

### 2024-01-24 outputs/paper/paper.qmd
Commiting all updates and work done to paper.qmd thus far.

@@
 The data presents a counterintuitive narrative, revealing that the University of Toronto, a Canadian fortress of education and innovation, experiences higher assault rates on average compared to its neighbouring communities.
-This finding challenges common perceptions of campus environments as insulated and safe.
-People often believe that a university is a safe bubble, away from the problems of the outside world, especially at a prestigious university focused on cutting edge learning and progress.
+This finding challenges common perceptions of campus environments as "insulated" and safe.
+People often believe that a university is a safe bubble, away from the problems of the outside world, especially at a prestigious university focused on cutting edge learning and societal progress.
 However, that is not what the data portrays.
@@
 ```
+The "University" neighborhood is the neighbourhood predominantly occupied by the University of Toronto campus.
+The steady increase in assault rates in the "University" neighborhood since 2020, as shown (\@ref(fig-assaults_per_year_per_neighbourhoods)), is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
+Unlike the surrounding areas, which display relatively stable or minimal fluctuating rates over the same period, the "University" exhibits a distinct and continuous rise.
+The next steepest rise in assault rates from 2020-2023 is the "Annex", which is another recognized University of Toronto Neighbourhood.
+These patterns within within the neighbourhoods "Annex" and "University" should raise concern and prompt conversation about assault rates on the University of Toronto campus.
 
+This divergence in trends between the "University" neighborhood and its neighboring areas suggests that the problem may be localized and thus could require targeted interventions by the university.
+These interventions could range from increasing campus security, bolstering support services for students, to implementing community engagement initiatives aimed at preventing assaults.
+
+There are potential measurment errors in the dataset used for this analysis.
+The dataset might not capture all incidents, as not all assaults are reported.
+Factors such as the victim's willingness to report, trust in campus authorities, and awareness of reporting channels can influence the data.
+Perhaps trust in campus authorities and awareness of the importance to report is strong on campus and therefore more assaults are reported in these neighbourhoods.
+
+There are several limitations to this analysis.
+Firstly, this analysis relies solely on a single metric (assault rates), which provides a narrow view of campus safety.
+Further, while assault rates does quantifies the incidents occurring, it does not account for the severity or nature of the assaults (does not break the data down by type of assault or whether it was a chargable offense).
+The term 'assault' encompasses a broad range of actions.
+Without information on the type of assault, it is difficult to provide insight into policy change, social action, and campus safety.
 # References

### 2024-01-24 outputs/paper/paper.qmd
Updating and commiting some spellcheck changes to paper.qmd

@@
 
-Although the Univerdity of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and voilence, there is very little data and literature that exists on the matter.
-This paper aids to contribute to the investigation of violence in and around the Universities district.
+Although the University of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and violence, there is very little data and literature that exists on the matter.
+This paper aids to contribute to the investigation of assault violence in and around the University of Toronto.
 
@@
 To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion.
-In the Data section, I discuss the nature of the spreadsheet obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
+In the Data section, I discuss the nature of the dataset obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
 The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight.
@@
 The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates].
-Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
+Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `KableExtra` [@rKableExtra], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
 
@@
 Data also includes the crime rate per 100,000 population calculated using the population estimates provided by Environics Analytics.
-Comparing to assault count, assault rate provides a fairer comparison of the crime over time by taking into account the change in population in the region.
+Compared to assault count, assault rate provides a fairer comparison of the crime over time by taking into account the change in population in the region.
 The data was last refreshed on Jan 11, 2024 and is refreshed annually.
+
 I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year.
@@
 The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.
+
+```{r}
+THIS IS FOR MY PHOTO
+```
 
@@
 Without information on the type of assault, it is difficult to provide insight into policy change, social action, and campus safety.
+
 # References

### 2024-01-24 scripts/00-simulate_data.R
Final commits to scripts.

@@ -46,7 +46,7 @@ simulated_data <- tibble(
 )
 
 # Summarize data for the two periods
-simulated_data <- simulated_data %>% 
+simulated_data <- simulated_data |> 
   mutate(
     Total_Assault_2014_2018 = ASSAULT_2014 + ASSAULT_2015 + ASSAULT_2016 + ASSAULT_2017 + ASSAULT_2018,
     Total_Assault_2019_2023 = ASSAULT_2019 + ASSAULT_2020 + ASSAULT_2021 + ASSAULT_2022 + ASSAULT_2023

### 2024-01-24 scripts/01-download_data.R

@@ -27,7 +27,7 @@ resources <- list_package_resources("neighbourhood-crime-rates")
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data <- filter(datastore_resources, row_number()==1) |> get_resource()
 data
 
 # Write the data to a CSV file

### 2024-01-24 scripts/02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw data to prepare for data anlysis 
+# Purpose: Cleans the raw data to create outputs/data/cleaned_data.csv
 # Author: Bella MacLean
 # Date: 21 January 2024
 # Contact: bella.maclean@mail.utoronto.ca

### 2024-01-24 scripts/02.1-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw data to prepare for data anlysis 
+# Purpose: Cleans the raw data to create outputs/data/analysis_data.csv
 # Author: Bella MacLean
 # Date: 21 January 2024
 # Contact: bella.maclean@mail.utoronto.ca

### 2024-01-24 scripts/02.2_data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw data to prepare for data anlysis 
+# Purpose: Cleans the raw data to create outputs/data/discussion_data.csv
 # Author: Bella MacLean
 # Date: 21 January 2024
 # Contact: bella.maclean@mail.utoronto.ca

### 2024-01-24 outputs/paper/paper.qmd
Title and abstract change.

@@
 ---
-title: "Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
+title: "How Safe is Campas Really? Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
 author: "Bella MacLean"
@@
 These 13 neighbourhoods encompass and surround the University of Toronto.
-This papers seeks to determine a correlation between the university's presence and assault rates in its neighbouring areas, providing insights that may inform campus safety strategies and urban planning."
+This papers seeks to determine a correlation between the university's presence and assault rates, providing insights that may inform campus safety strategies and urban planning."
 format: pdf

### 2024-01-24 outputs/paper/paper.qmd
made a couple naming changes.

@@
 
-long_data <- cleaned_data |>
+assault_data <- cleaned_data |>
   gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
@@
   ) +
-  ylim(0, 5000) # Adjust these values based on your data's range
+  ylim(0, 3000) # Adjust these values based on your data's range
 
@@
 
-long_data <- surrounding_neighbourhoods |>
+asst_data <- surrounding_neighbourhoods |>
   gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
@@
 
-# Your data preparation code
+# My data prep
 long_data <- all_neighbourhoods |>
@@
 
-# Plot the graph with facets and without a legend
+# Plot the graph with facets and without a legend because legend was to big and messy and every facet is already labelled nicely
 ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
@@
 
-There are potential measurment errors in the dataset used for this analysis.
+There are potential measurement errors in the dataset used for this analysis.
 The dataset might not capture all incidents, as not all assaults are reported.

### 2024-01-24 outputs/paper/paper.qmd
Added some small formatting changes.

@@
 
+\newpage
+
 ##Results
@@
 ```
+
+\newpage
+
 #Discussion
@@
 
+\newpage
+
 # References

### 2024-01-24 outputs/paper/paper.qmd
Added my maps.

@@
 
-```{r}
-THIS IS FOR MY PHOTO
-```
-
 ###University of Toronto Neighbourhoods
@@
 The next crucial step was to narrow down the neighbourhoods to the four aforementioned areas central to the University of Toronto.
-These neighbourhoods are 4 of the 158 legitamate neighbourhoods recognized by the city of Toronto and can be seen in figure (PICTURE REF> HERE).
-This selective focus was vital to the study's goal of examining crime trends in university-adjacent areas (See Table \@ref(tbl-cleaned_uoft_neighbourhood_data)).
+These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in (@ref(fig-uoft_neighbourhood_map)).
+This selective focus was vital to the study's goal of examining assault trends in university-adjacent areas (See Table \@ref(tbl-cleaned_uoft_neighbourhood_data)).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-uoft_neighbourhood_map
+#| fig-cap: Map of UofT Neighbourhoods
+
+knitr::include_graphics("/cloud/project/inputs/sketches/uoft_neighbourhoods.png")
+```
 
@@
 This was done to understand the peripheral impact and the broader context of assault rates around the university.
-This secondary cleaning phase involved identifying and including adjacent neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See Table \@ref(tbl-cleaned_surrounding_neighbourhood_data)).
-These neighbourhoods make up 10 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in figure (PICTURE REF> HERE).
+This secondary cleaning phase involved identifying and including surrounding neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See Table \@ref(tbl-cleaned_surrounding_neighbourhood_data)).
+These neighbourhoods make up 9 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in (@ref(fig-surrounding_neighbourhood_map)).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-surrounding_neighbourhood_map
+#| fig-cap: Map of All 13 Neighbourhoods
+
+knitr::include_graphics("/cloud/project/inputs/sketches/surrounding_neighbourhoods.png")
+```
 

### 2024-01-24 outputs/paper/paper.qmd
Added some comments and my conclusion.

@@
 
+# Adding my map of UofT neighbourhoods
 knitr::include_graphics("/cloud/project/inputs/sketches/uoft_neighbourhoods.png")
@@
 
+# Adding my map of surrounding neighbourhoods
 knitr::include_graphics("/cloud/project/inputs/sketches/surrounding_neighbourhoods.png")
@@
   ) +
-  ylim(0, 3000) # Adjust these values based on your data's range
+  ylim(0, 3000) # Adjust these values based on my data's range
 
@@
 
+# Conclusion
+This analysis of assault rates from 2014 to 2023 in and around the University of Toronto offers a unique view of campus safety in raletion to assault rates and its surrounding environment.
+The significant findings challenge the commonly held belief that university campuses, especially prestigious ones like the University of Toronto, are insulated and often safer then surrounding areas.
+The data shows a notable increase in assault rates in the "University" and "Annex" neighborhoods, areas predominantly associated with the University of Toronto.
+This upward trend, particularly since 2020, emphasizes the necessity for a reassessment of the current safety measures and policies in place on campus.
+The contrasting stability or minor fluctuations in assault rates in the surrounding neighborhoods further accentuates the specific challenges faced within the university's immediate vicinity.
+This analysis contributes to the dialogue around campus safety.
+The findings suggest that the University of Toronto might benefit from increased targeted interventions in relation to assault prevention and protection.
+
 \newpage

### 2024-01-24 outputs/paper/paper.qmd
saving some formatting changes
Final commits before peer review.

@@
 library(stringr)
+library(kableExtra)
+library(bookdown)
 
@@
 
-## Introduction
+# Introduction
 
@@
 
-## Data
+# Data
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
 The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates].
-Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `KableExtra` [@rKableExtra], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
+Data was collected and analyzed using the statistical programming software R [@citeR], with additional support from `tidyverse` [@citetidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `KableExtra` [@rKableExtra], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
 
-###Source and Data Collection
+### Source and Data Collection
 This dataset includes the Crime Data by Neighbourhood.
@@
 
-###University of Toronto Neighbourhoods
+### University of Toronto Neighbourhoods
 The 'Neighbourhood Crime Rates' dataset was filtered to focus solely on assault rates.
@@
 The next crucial step was to narrow down the neighbourhoods to the four aforementioned areas central to the University of Toronto.
-These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in (@ref(fig-uoft_neighbourhood_map)).
-This selective focus was vital to the study's goal of examining assault trends in university-adjacent areas (See Table \@ref(tbl-cleaned_uoft_neighbourhood_data)).
+These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in Figure \@ref(fig-uoft_neighbourhood_map).
+This selective focus was vital to the study's goal of examining assault trends in university-adjacent areas.
+See Table \@ref(tbl-cleaned_uoft_neighbourhood_data).
 
-```{r}
+```{r fig-uoft_neighbourhood_map, echo=FALSE, out.width='100%', fig.align='center'}
 #| message: false
@@
 
-```{r}
+```{r tbl-cleaned_uoft_neighbourhood_data, table-fit, echo=FALSE, results='asis'}
 #| message: false
@@
 
-# Displaying the first 5 rows of the dataset
-head(cleaned_data, 5) |>
-  kable(
-    col.names = c("Neighbourhood", "Assault Rate 2014", "Assault Rate 2015", "Assault Rate 2016", "Assault Rate 2017", "Assault Rate 2018", "Assault Rate 2019", "Assault Rate 2020", "Assault Rate 2021", "Assault Rate 2022", "Assault Rate 2023"),
-    booktabs = TRUE
-  )
+# Select only the first four columns and the first 5 rows of the dataset
+selected_data_uoft <- head(uoft_neighbourhoods, 5)[, 1:4]
+
+# Displaying the table with only the first four columns
+kable(selected_data_uoft, 
+      booktabs = TRUE) |>
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))
+
 ```
 
-###University of Toronto Surrounding Areas
+### University of Toronto Surrounding Areas
 The scope was then expanded to encompass neighbourhoods surrounding the primary four that make up the University of Toronto.
@@
 
-```{r}
+```{r surr-map, echo=FALSE, out.width='100%', fig.align='center'}
 #| message: false
@@
 
-```{r}
+```{r table-fit, echo=FALSE, results='asis'}
 #| message: false
@@
 
-# Displaying the first 5 rows of the dataset
-head(surrounding_neighbourhoods, 5) |>
-  kable(
-    col.names = c("Neighbourhood", "Assault Rate 2014", "Assault Rate 2015", "Assault Rate 2016", "Assault Rate 2017", "Assault Rate 2018", "Assault Rate 2019", "Assault Rate 2020", "Assault Rate 2021", "Assault Rate 2022", "Assault Rate 2023"),
-    booktabs = TRUE
-  )
+# Select only the first four columns and the first 10 rows of the dataset
+selected_data <- head(surrounding_neighbourhoods, 10)[, 1:4]
+
+# Displaying the table with only the first four columns
+kable(selected_data, 
+      format = "latex", booktabs = TRUE) |>
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))
+
 ```
@@
 
-##Results
-###Assault Rates per University of Toronto Neighbourhood 
+# Results
+### Assault Rates per University of Toronto Neighbourhood 
 Over the past decade, University of Toronto's encompassing neighbourhoods have exhibited unique trends in assault rates.
@@
 (\@ref(fig-assaults_per_year_uoft)) shows the assault rates per year for each University of Toronto neighbourhood.
-```{r}
+```{r fig-assaults_per_year_uoft, message=FALSE, echo=FALSE, fig.cap="Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)"}
 #| message: false
@@
 
-assault_data <- cleaned_data |>
+# Reading the data
+uoft_neighbourhoods <- read.csv(here::here("outputs/data/cleaned_data.csv"))
+
+long_data <- uoft_neighbourhoods |>
   gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
-  mutate(Year = as.numeric(str_replace(Year, "ASSAULT_RATE_", "")))
+  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")))
 
@@
   labs(
-    title = "Assault Rates in University of Toronto 
-      Neighbourhoods (2014-2023)",
+    title = "   Assault Rates in University of Toronto's 
+Surrounding Neighbourhoods (2014-2023)",
     x = "Year",
@@
   ) +
-  ylim(0, 3000) # Adjust these values based on my data's range
+  ylim(0, 3000) +
+  scale_x_continuous(breaks = seq(from = 2014, to = 2023, by = 2)) # I added this line because it specifies the break points for the x-axis, which in this context represents the years.
+This line is telling ggplot2 to place axis ticks at every integer value between 2014 and 2023.
+I had to add this because the years were not showing up there was something wrong at the x-axis was coming up blank and therefore there was no datapoints in the graph
 
@@
 
-###Assault Rates per Neighbourhood in the Surrounding Areas
+### Assault Rates per Neighbourhood in the Surrounding Areas
 Assault rates in the neighbourhoods surrounding the University of Toronto Neighbourhoods reveals some distinct patterns and fluctuations in the data.
@@
 For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to (\@ref(fig-assaults_per_year_surounding_neighbourhoods)).
-```{r}
+``````{r fig-assaults_per_year_surounding_neighbourhoods, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Assault Rates in Surrounding Neighbourhoods"}
 #| message: false
@@
 
-asst_data <- surrounding_neighbourhoods |>
+long_data <- surrounding_neighbourhoods |>
   gather(key = "Year", value = "Assault_Rate", -Neighbourhood) |>
   mutate(Year = as.numeric(str_extract(Year, "\\d{4}")))
-
-# Checking for NA values in Year because when I was building my graph orignally the output was a graph with nothing plotted on it even though everything else looked good the Year column is producing NA values
-sum(is.na(long_data$Year))
 
@@
 
-#Discussion
+# Discussion
 The data allows for a comparative analysis of assault rates across various neighborhoods in and around the University of Toronto from 2014 to 2023.
@@
   labs(
-    title = "Assault Rates in University of Toronto's Surrounding Neighbourhoods (2014-2023)",
+    title = "Assault Rates per year Across All Neighbourhoods From 2014-2023",
     x = "Year",
@@
 The next steepest rise in assault rates from 2020-2023 is the "Annex", which is another recognized University of Toronto Neighbourhood.
-These patterns within within the neighbourhoods "Annex" and "University" should raise concern and prompt conversation about assault rates on the University of Toronto campus.
+These patterns within the neighbourhoods "Annex" and "University" should raise concern and prompt conversation about assault rates on the University of Toronto campus.
 

### 2024-01-24 outputs/paper/paper.qmd
Final commits before peer review.

@@
 These 13 neighbourhoods encompass and surround the University of Toronto.
-This papers seeks to determine a correlation between the university's presence and assault rates, providing insights that may inform campus safety strategies and urban planning."
+This papers seeks to determine a correlation between the university's presence and assault rates, providing insights that may inform campus safety strategies and policy."
 format: pdf

### 2024-01-24 scripts/02-data_cleaning.R
Peer reviews changes include changing the decimal places to 2 and fixing a spelling error in my title.

@@ -31,7 +31,19 @@ assault_rate_data <- raw_data |>
          `Assault Rate 2020` = ASSAULT_RATE_2020,
          `Assault Rate 2021` = ASSAULT_RATE_2021,
          `Assault Rate 2022` = ASSAULT_RATE_2022,
-         `Assault Rate 2023` = ASSAULT_RATE_2023)
+         `Assault Rate 2023` = ASSAULT_RATE_2023) |>
+  mutate(
+    `Assault Rate 2014` = round(`Assault Rate 2014`, 2),
+    `Assault Rate 2015` = round(`Assault Rate 2015`, 2),
+    `Assault Rate 2016` = round(`Assault Rate 2016`, 2),
+    `Assault Rate 2017` = round(`Assault Rate 2017`, 2),
+    `Assault Rate 2018` = round(`Assault Rate 2018`, 2),
+    `Assault Rate 2019` = round(`Assault Rate 2019`, 2),
+    `Assault Rate 2020` = round(`Assault Rate 2020`, 2),
+    `Assault Rate 2021` = round(`Assault Rate 2021`, 2),
+    `Assault Rate 2022` = round(`Assault Rate 2022`, 2),
+    `Assault Rate 2023` = round(`Assault Rate 2023`, 2)
+  )
 
 # Filter rows for specific neighborhoods
 uoft_neighbourhoods <- assault_rate_data |>

### 2024-01-24 scripts/02.1-data_cleaning.R

@@ -31,7 +31,20 @@ assault_rate_data <- raw_data |>
          `Assault Rate 2020` = ASSAULT_RATE_2020,
          `Assault Rate 2021` = ASSAULT_RATE_2021,
          `Assault Rate 2022` = ASSAULT_RATE_2022,
-         `Assault Rate 2023` = ASSAULT_RATE_2023)
+         `Assault Rate 2023` = ASSAULT_RATE_2023) |>
+  mutate(
+    `Assault Rate 2014` = round(`Assault Rate 2014`, 2),
+    `Assault Rate 2015` = round(`Assault Rate 2015`, 2),
+    `Assault Rate 2016` = round(`Assault Rate 2016`, 2),
+    `Assault Rate 2017` = round(`Assault Rate 2017`, 2),
+    `Assault Rate 2018` = round(`Assault Rate 2018`, 2),
+    `Assault Rate 2019` = round(`Assault Rate 2019`, 2),
+    `Assault Rate 2020` = round(`Assault Rate 2020`, 2),
+    `Assault Rate 2021` = round(`Assault Rate 2021`, 2),
+    `Assault Rate 2022` = round(`Assault Rate 2022`, 2),
+    `Assault Rate 2023` = round(`Assault Rate 2023`, 2)
+  )
+
 
 # Filter rows for specific neighborhoods
 surrounding_neighbourhoods <- assault_rate_data |>

### 2024-01-25 outputs/paper/paper.qmd
Peer reviews changes include changing the decimal places to 2 and fixing a spelling error in my title.
This commit includes a lot of formatting changes. Including all cross references that are now working to be expected. Changed some graph and table formatting.

@@
 ---
-title: "How Safe is Campas Really? Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
+title: "How Safe is Campus Really? Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
 author: "Bella MacLean"
-thanks: "Code and data from this analysis are available at: https://github.com/Bellamaclean7/UofT_Neighbourhood_Assaults/tree/main"
+thanks: "Code and data from this analysis are available at: https://github.com/Bellamaclean7/UofT_Neighbourhood_Assault_Data"
 date: "today"
@@
 The next crucial step was to narrow down the neighbourhoods to the four aforementioned areas central to the University of Toronto.
-These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in Figure \@ref(fig-uoft_neighbourhood_map).
+These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in @fig-uoft_neighbourhood_map.
 This selective focus was vital to the study's goal of examining assault trends in university-adjacent areas.
-See Table \@ref(tbl-cleaned_uoft_neighbourhood_data).
+See @tbl-cleaned_uoft_neighbourhood_data.
 
 ```{r fig-uoft_neighbourhood_map, echo=FALSE, out.width='100%', fig.align='center'}
-#| message: false
+#| warning: false
 #| echo: false
@@
 This was done to understand the peripheral impact and the broader context of assault rates around the university.
-This secondary cleaning phase involved identifying and including surrounding neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See Table \@ref(tbl-cleaned_surrounding_neighbourhood_data)).
-These neighbourhoods make up 9 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in (@ref(fig-surrounding_neighbourhood_map)).
+This secondary cleaning phase involved identifying and including surrounding neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See @tbl-cleaned_surrounding_neighbourhood_data).
+These neighbourhoods make up 9 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in @fig-surrounding_neighbourhood_map.
 
-```{r surr-map, echo=FALSE, out.width='100%', fig.align='center'}
+```{r fig-surrounding_neighbourhood_map, echo=FALSE, out.width='100%', fig.align='center'}
 #| message: false
@@
 The Annex also shows an increase from 2020 on, though not as steep as University.
-(\@ref(fig-assaults_per_year_uoft)) shows the assault rates per year for each University of Toronto neighbourhood.
+@fig-assaults_per_year_uoft shows the assault rates per year for each University of Toronto neighbourhood.
 ```{r fig-assaults_per_year_uoft, message=FALSE, echo=FALSE, fig.cap="Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)"}
@@
 Clair", "Casa Loma", "Palmerston-Little Italy", and "Wychwood".
-For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to (\@ref(fig-assaults_per_year_surounding_neighbourhoods)).
+For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to @fig-assaults_per_year_surounding_neighbourhoods.
 ``````{r fig-assaults_per_year_surounding_neighbourhoods, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Assault Rates in Surrounding Neighbourhoods"}
@@
 People often believe that a university is a safe bubble, away from the problems of the outside world, especially at a prestigious university focused on cutting edge learning and societal progress.
-However, that is not what the data portrays.
-(See\@ref(fig-assaults_per_year_per_neighbourhoods))
+However, that is not what the data portrays (See @fig-assaults_per_year_per_neighbourhoods).
 
@@
     legend.position = "none", # This line removes the legend
-    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
+    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+    plot.title = element_text(size = 25, hjust = 0.5), # Adjusting title
+    axis.title.x = element_text(size = 16), # Adjusting x-axis title
+    axis.title.y = element_text(size = 16) # Adjusting y-axis title
   )
@@
 The "University" neighborhood is the neighbourhood predominantly occupied by the University of Toronto campus.
-The steady increase in assault rates in the "University" neighborhood since 2020, as shown (\@ref(fig-assaults_per_year_per_neighbourhoods)), is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
+The steady increase in assault rates in the "University" neighborhood since 2020, as shown in @fig-assaults_per_year_per_neighbourhoods, is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
 Unlike the surrounding areas, which display relatively stable or minimal fluctuating rates over the same period, the "University" exhibits a distinct and continuous rise.

### 2024-01-25 scripts/03-test_data.R
Added all my testing and saved to test_data.R

@@ -1,15 +1,42 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Ran tests on uoft_neighbourhoods, surrounding_neighbourhoods, and all_neighbourhoods to ensure the data is as expected
+# Author: Bella MacLean
+# Date: 25 January 2024
+# Contact: bella.maclean@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: none
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Test data ####
+# Read in cleaned data #
+uoft_neighbourhoods <- read.csv(here::here("outputs/data/cleaned_data.csv"))
+show_col_types = FALSE
+
+# Read in cleaned data #
+surrounding_neighbourhoods <- read.csv(here::here("outputs/data/analysis_data.csv"))
+show_col_types = FALSE
+
+# Read in cleaned data #
+all_neighbourhoods <- read.csv(here::here("outputs/data/discussion_data.csv"))
+show_col_types = FALSE
+
+# Checking for any duplicated rows in the dataframes 
+any(duplicated(uoft_neighbourhoods))
+any(duplicated(surrounding_neighbourhoods))
+any(duplicated(all_neighbourhoods))
+
+# Checking to ensure the number of rows is as expected
+nrow(uoft_neighbourhoods) # I want there to be 4 rows
+nrow(surrounding_neighbourhoods) # I want there to be 9 rows
+nrow(all_neighbourhoods) # I want there to be 13 rows
+
+# Checking to ensure there are no missing values in the dataframes
+sum(is.na(uoft_neighbourhoods))
+sum(is.na(surrounding_neighbourhoods))
+sum(is.na(all_neighbourhoods))
+
+# Checking to ensure the dataframe holds only the unique neighbourhoods I want
+unique(uoft_neighbourhoods$Neighbourhood)
+unique(surrounding_neighbourhoods)
+unique(all_neighbourhoods)

### 2024-01-25 scripts/03-test_data.R
Had a small error in one of my tests

@@ -38,5 +38,5 @@ sum(is.na(all_neighbourhoods))
 
 # Checking to ensure the dataframe holds only the unique neighbourhoods I want
 unique(uoft_neighbourhoods$Neighbourhood)
-unique(surrounding_neighbourhoods)
-unique(all_neighbourhoods)
+unique(surrounding_neighbourhoods$Neighbourhood)
+unique(all_neighbourhoods$Neighbourhood)

### 2024-01-26 outputs/paper/paper.qmd
Added a reference. Worked on discussion section. Updated title. Made some formatting changes. Checked spelling.

@@
 ---
-title: "How Safe is Campus Really? Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
+title: "Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
 author: "Bella MacLean"
@@
 ```
+\newpage
 
@@
 Over 117,000 students from various institutions, including over 20,000 from U of T, participated in the survey.
-The overview of the survey released indicated that 63 per cent of students province-wide, and 59 per cent of students at U of T, reported at least one incident of sexual harassment since the beginning of the 2017-2018 academic year [@citeUofTSexualViolenceData].
+The overview of the survey released indicated that 63 per cent of students province-wide, and 59 per cent of students at U of T, reported at least one incident of sexual assault since the beginning of the 2017-2018 academic year [@citeUofTSexualViolenceData].
 
@@
 This paper aids to contribute to the investigation of assault violence in and around the University of Toronto.
+
+During the literature review phase of this analysis, it became evident that the majority of available data and academic discourse concerning assault crime on campus predominantly focused on sexual assault and harassment.
+This imbalance in data representation potentially masks a full understanding of the prevalence and nature of other types of assault that may be taking place.
+This lack of data on various forms of assault could hinder the universities ability to grasp the complete spectrum of safety issues faced by those on campus.
 
@@
 
-\newpage
-
 # Data
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
-The data set used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates].
+The dataset used is entitled 'Neighbourhood Crime Rates' [@citeTorontoCrimeRates].
 Data was collected and analyzed using the statistical programming software R [@citeR], with additional support from `tidyverse` [@citetidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `KableExtra` [@rKableExtra], `tibble` [@rTibble], `janitor` [@rJanitor], and `knitr` [@rknitr].
@@
 The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.
+University of Toronto neighbourhoods include: "Bay-Cloverhill", "Annex", "University", and "Kensington-Chinatown".
+Surrounding neighbourhoods include: "Dovercourt Village", "Church-Wellesley", "Wellington Place", "West Queen West", "Yonge-St.
+Clair", "Casa Loma", "Wychwood", "Palmerston-Little Italy", and "Trinity-Bellwoods".
 
@@
 This entailed discarding unrelated crime statistics and zeroing in on annual assault rate data per neighbourhood.
-The next crucial step was to narrow down the neighbourhoods to the four aforementioned areas central to the University of Toronto.
+The next crucial step was to narrow down the neighbourhoods to the 4 aforementioned areas central to the University of Toronto.
 These neighbourhoods are 4 of the 158 legitimate neighbourhoods recognized by the city of Toronto and can be seen in @fig-uoft_neighbourhood_map.
@@
 ### University of Toronto Surrounding Areas
-The scope was then expanded to encompass neighbourhoods surrounding the primary four that make up the University of Toronto.
-This was done to understand the peripheral impact and the broader context of assault rates around the university.
+The scope was then expanded to encompass neighbourhoods surrounding the primary 4 that make up the University of Toronto.
+This was done to understand the peripheral impact and the broader context of assault rates in and around the university.
 This secondary cleaning phase involved identifying and including surrounding neighbourhoods, thereby painting a more holistic picture of the university's neighbourhood safety landscape (See @tbl-cleaned_surrounding_neighbourhood_data).
@@
 
-\newpage
-
 # Results
@@
 #| label: fig-assaults_per_year_uoft
-#| fig-cap: Assault Rates per year in University of Toronto Neighbourhoods (2014-2023)
+#| fig-cap: Assault Rates in University of Toronto Neighbourhoods (2014-2023)
 
@@
 
+# This first part of the code takes "assault_rate_[year]" and consolidates them into two columns: one for the year (key = "Year") and one for the corresponding assault rates (value = "Assault_Rate") - I did this to improve data visualization using line graphs
 long_data <- uoft_neighbourhoods |>
@@
 
-# Plot the graph
+# Plotting the graph
 ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
@@
 ### Assault Rates per Neighbourhood in the Surrounding Areas
-Assault rates in the neighbourhoods surrounding the University of Toronto Neighbourhoods reveals some distinct patterns and fluctuations in the data.
+Assault rates in neighbourhoods surrounding the University of Toronto reveals some distinct patterns and fluctuations.
 Several neighbourhoods exhibit fairly stable trends with minor year-to-year changes including "Yonge-St.Clair", "Wychwood", "West Queen West, "Palmerston-Little Italy", "Case Loma" and "Trinity-Bellwoods".
@@
 "Church-Wellesley" demonstrate quite a sporadic trend with several jumps and falls.
-From 2014-2020 "Wellington Place" shows a steep downward trend, rising from 2020-2022, however, then showing a steep decline again from 2022-2023.
+From 2014-2020 "Wellington Place" shows a steep downward trend, rising from 2020-2022, however, then again showing a steep decline.
 Several neighbourhoods exhibit quite low assault rates over the entire decade including "Yonge-St.
@@
 #| label: fig-assaults_per_year_surounding_neighbourhoods
-#| fig-cap: Assault Rates in Surrounding Neighbourhoods
+#| fig-cap: Assault Rates in Surrounding Neighbourhoods (2014-2023)
 
+# This first part of the code takes "assault_rate_[year]" and consolidates them into two columns: one for the year (key = "Year") and one for the corresponding assault rates (value = "Assault_Rate") - I did this to improve data visualization using line graphs
 long_data <- surrounding_neighbourhoods |>
@@
 
-\newpage
-
 # Discussion
 The data allows for a comparative analysis of assault rates across various neighborhoods in and around the University of Toronto from 2014 to 2023.
-By comparing the assault trends in the university's immediate vicinity with those of the surrounding neighborhoods, I aim to uncover underlying patterns, draw attention to areas of concern, and spark discussions around potential causative factors.
+By comparing the assault trends in the university's immediate vicinity with those of the surrounding neighborhoods, I aim to uncover underlying patterns, draw attention to areas of concern, and spark discussions around potential preventative measures.
 
@@
 #| label: fig-assaults_per_year_per_neighbourhoods
-#| fig-cap: Assault Rates per year Across All Neighbourhoods (2014-2023)
+#| fig-cap: Assault Rates Across All Neighbourhoods (2014-2023)
 
@@
 
-# Plot the graph with facets and without a legend because legend was to big and messy and every facet is already labelled nicely
+# Plot the graph with facets and without a legend because legend was to big and messy and every subplot is already labelled nicely with neighbourhood name
 ggplot(long_data, aes(x = Year, y = Assault_Rate, group = Neighbourhood, color = Neighbourhood)) +
@@
     y = "Assault Rate"
-  ) +
+  ) + # this next part formats the x-axis scale to include every year
   scale_x_continuous(breaks = seq(from = 2014, to = 2023, by = 1)) +
@@
 ```
-The "University" neighborhood is the neighbourhood predominantly occupied by the University of Toronto campus.
+"University" is the neighbourhood predominantly occupied by the University of Toronto campus.
 The steady increase in assault rates in the "University" neighborhood since 2020, as shown in @fig-assaults_per_year_per_neighbourhoods, is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
-Unlike the surrounding areas, which display relatively stable or minimal fluctuating rates over the same period, the "University" exhibits a distinct and continuous rise.
-The next steepest rise in assault rates from 2020-2023 is the "Annex", which is another recognized University of Toronto Neighbourhood.
-These patterns within the neighbourhoods "Annex" and "University" should raise concern and prompt conversation about assault rates on the University of Toronto campus.
+Unlike the surrounding areas, which display relatively stable or minimal fluctuating rates over the same period, "University" exhibits a distinct and continuous rise.
+The next steepest rise in assault rates from 2020-2023 is "Annex", which is another recognized University of Toronto Neighbourhood.
+Additionally, assault rates in "Kensington-Chinatown" are most consistent and highest overall across the entirety of the decade then all surrounding neighnourhoods.
+These patterns within the University of Toronto neighbourhoods should raise concern and prompt conversation about assault rates on the University of Toronto campus.
 
-This divergence in trends between the "University" neighborhood and its neighboring areas suggests that the problem may be localized and thus could require targeted interventions by the university.
-These interventions could range from increasing campus security, bolstering support services for students, to implementing community engagement initiatives aimed at preventing assaults.
+This divergence in trends between the "University" neighborhood and its neighboring areas suggests that the problem may be localized and thus require targeted interventions by the university.
+These interventions could range from increasing campus security, bolstering support services for students, and implementing community engagement initiatives aimed at preventing assaults of all types.
 
+The 2022 Statistical Overview of Reportable Incidents at University of Toronto - St.
+George, released by the University of Toronto Campus Safety Special Constable Service showed that general assault counts are 3 times higher than sexual assault counts [@citeUofTCampusSafetyReport].
+However, emphasis is still put on sexual assault prevention almost exclusively in all popular literature.
+A major flaw to this important yearly circulated document is that the report does not discuss or capture campus boundaries in relation to Special Constable Services.
+
+Given that the data indicates higher assault rates in the University of Toronto neighbourhoods compared to surrounding areas, and the predominance of discourse strictly referencing sexual assault in the literature, it's imperative that the university works expose data and spread awareness concerning all types of assault present within campus neighbourhoods.
+It is also crucial that the University of Toronto feel a sense of responsibility to geographical pockets immediately surrounding the university with high off-campus student populations.
+Future research should aim to record assault data by type, severity, and location, providing the most comprehensive understanding of the issue possible.
+
+### Limitations in the Data
 There are potential measurement errors in the dataset used for this analysis.
 The dataset might not capture all incidents, as not all assaults are reported.
-Factors such as the victim's willingness to report, trust in campus authorities, and awareness of reporting channels can influence the data.
+Factors such as the victim's willingness to report, trust in authorities and institutions, and awareness of reporting channels can influence the data.
 Perhaps trust in campus authorities and awareness of the importance to report is strong on campus and therefore more assaults are reported in these neighbourhoods.
@@
 Firstly, this analysis relies solely on a single metric (assault rates), which provides a narrow view of campus safety.
-Further, while assault rates does quantifies the incidents occurring, it does not account for the severity or nature of the assaults (does not break the data down by type of assault or whether it was a chargable offense).
+Further, while assault rates does quantify the incidents occurring, it does not account for the severity or nature of the assaults (does not break the data down by type of assault or whether it was a chargeable offense).
 The term 'assault' encompasses a broad range of actions.
-Without information on the type of assault, it is difficult to provide insight into policy change, social action, and campus safety.
+Without information on the type of assault, it is difficult to provide targeted insight into policy change, social action, and campus safety.
 

### 2024-01-26 scripts/00-simulate_data.R
Cleaned up and revised some of my comments in my simulated data and cleaned data

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate the Neighbourhood Crime Rates data set but refining the 158 neighbourhoods  
+# Purpose: Simulate the Neighbourhood Crime Rates data set by refining the 158 neighbourhoods  
 # Author: Bella MacLean
 # Date: Saturday January 20, 2024
 # Contact: bella.maclean@mail.utoronto.ca

### 2024-01-26 scripts/02.1-data_cleaning.R

@@ -14,7 +14,7 @@ library(readr)
 raw_data <- read_csv("inputs/data/raw_data.csv")
 View(raw_data)
 
-# Select the specified columns and rename AREA_NAME
+# Select the specified columns and rename AREA_NAME and round each assault rate to 2 decimal places
 assault_rate_data <- raw_data |>
   select(AREA_NAME, 
          ASSAULT_RATE_2014, ASSAULT_RATE_2015, ASSAULT_RATE_2016, 

### 2024-01-26 outputs/paper/paper.qmd
Changed my title.

@@
 ---
-title: "Assault Rate Trends Within and Around the University of Toronto: A Ten-Year Analysis"
+title: "Assault Crime at the University of Toronto: Higher Than in Surrounding Neighbourhoods?"
+subtitle: "Exploring a Decade of Assault Rate Trends in and Around the University"
 author: "Bella MacLean"

### 2024-01-26 outputs/paper/paper.qmd
Small change to abstract.

@@
 These 13 neighbourhoods encompass and surround the University of Toronto.
-This papers seeks to determine a correlation between the university's presence and assault rates, providing insights that may inform campus safety strategies and policy."
+This paper reveals a correlation between the university's presence and higher assault rates, providing insights that may inform campus safety strategies and policy."
 format: pdf

### 2024-01-26 outputs/paper/paper.qmd
Changed a few thing in the introduction including adding an sub heading for analysis structure. Change a couple other sub heading to make them all match vocabulary better.

@@
 This paper examines assault rates from 2014 to 2023 in 13 of Toronto's 158 recignized neighbourhoods, using data from OpenDataToronto [@rOpenDataToronto], to understand the safety landscape of the University of Toronto in relation to assault rates over a ten-year period.
+This analysis sets out to expose higher and rising assault rates within University of Toronto Neighbourhoods compared to surrounding neighbourhoods.
 
@@
 Although the University of Toronto emphasizes its ongoing commitment to student safety, the improvement of services, and raising awareness of assault and violence, there is very little data and literature that exists on the matter.
-This paper aids to contribute to the investigation of assault violence in and around the University of Toronto.
-
 During the literature review phase of this analysis, it became evident that the majority of available data and academic discourse concerning assault crime on campus predominantly focused on sexual assault and harassment.
@@
 
+### Analysis Structure 
 To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion.
@@
 
-### University of Toronto Surrounding Areas
+### University of Toronto Surrounding Neighbourhoods
 The scope was then expanded to encompass neighbourhoods surrounding the primary 4 that make up the University of Toronto.
@@
 
-### Assault Rates per Neighbourhood in the Surrounding Areas
+### Assault Rates per Surrounding Neighbourhood
 Assault rates in neighbourhoods surrounding the University of Toronto reveals some distinct patterns and fluctuations.
@@
 ```
+
+\newpage
 
@@
 This upward trend, particularly since 2020, emphasizes the necessity for a reassessment of the current safety measures and policies in place on campus.
-The contrasting stability or minor fluctuations in assault rates in the surrounding neighborhoods further accentuates the specific challenges faced within the university's immediate vicinity.
+The contrasting stability and/or minor fluctuations present in assault rates in the surrounding neighborhoods further accentuates the specific challenges faced within the university's immediate vicinity.
 This analysis contributes to the dialogue around campus safety.
-The findings suggest that the University of Toronto might benefit from increased targeted interventions in relation to assault prevention and protection.
+The findings suggest that the University of Toronto might benefit from increased targeted interventions in relation to assault prevention, as well as expanding resources and services to geographical pockets immediately surrounding the university with high levels of student presence.
 

### 2024-01-26 outputs/paper/paper.qmd
test.

@@
 
-### Analysis Structure 
 To examine the rate of assaults within and around the University of Toronto from 2014-2023, this paper has been organized into the following sections: Data, Results, Discussion, and Conclusion.

### 2024-01-26 outputs/paper/paper.qmd
Changed a few thing in the introduction including adding an sub heading for analysis structure. Change a couple other sub heading to make them all match vocabulary better.
Final edits.

--- 2024-01-26 outputs/paper/paper.qmd
@@
 Toronto is a city known for its vibrant culture and diverse neighborhoods and is home to one of Canada's most prestigious educational institutions – the University of Toronto.
-The University of Toronto encompasses and shapes the nearby areas of Bay-Cloverhill, The Annex, University district, and Kensington-Chinatown.
-These neighbourhoods are home to both an abundance of UofT academic and student buildings and a large student population.
-This paper examines assault rates from 2014 to 2023 in 13 of Toronto's 158 recignized neighbourhoods, using data from OpenDataToronto [@rOpenDataToronto], to understand the safety landscape of the University of Toronto in relation to assault rates over a ten-year period.
-This analysis sets out to expose higher and rising assault rates within University of Toronto Neighbourhoods compared to surrounding neighbourhoods.
+The University of Toronto encompasses and shapes the nearby areas of Bay-Cloverhill, The Annex, University District, and Kensington-Chinatown.
+These neighbourhoods are home to both an abundance of University of Toronto academic and student buildings and a large student population.
+This paper examines assault rates from 2014 to 2023 in 13 of Toronto's 158 recognized neighbourhoods, using data from OpenDataToronto [@rOpenDataToronto], to understand the safety landscape of the University of Toronto in relation to assault rates over a ten-year period.
+This analysis sets out to expose higher and rising assault rates within University of Toronto neighbourhoods compared to surrounding neighbourhoods.
 
 In February of 2020, the University of Toronto released data from a two-year-old province-wide, online student survey on sexual harassment and sexual assault, saying it will paint a more detailed picture of the issues and help improve supports on campus [@citeUofTSexualViolenceData].
-Originally collected as part of the 2018 Student Voices on Sexual Violence Survey, the data was then released by the Ontario government to all universities and colleges province wide.
+Originally collected as part of the 2018 Student Voices on Sexual Violence Survey, the data was then released by the Ontario government to all universities and colleges province-wide.
 Over 117,000 students from various institutions, including over 20,000 from U of T, participated in the survey.
@@
 During the literature review phase of this analysis, it became evident that the majority of available data and academic discourse concerning assault crime on campus predominantly focused on sexual assault and harassment.
-This imbalance in data representation potentially masks a full understanding of the prevalence and nature of other types of assault that may be taking place.
-This lack of data on various forms of assault could hinder the universities ability to grasp the complete spectrum of safety issues faced by those on campus.
+The current focus being mainly on one kind of assault (sexual assault) could be obscuring the true extent and variety of other assaults occurring.
+This shortage of data and discourse could be hindering the university from fully understanding the range of safety challenges that exist on campus and therefore implementing the appropriate safety strategies and policy.
 
@@
 I conducted the first step of basic data cleaning to eliminate all counts and rates not applicable to this paper, so all I was left with was assault rates per neighbourhood for each year.
-I then refined the nieghbourhoods to include only those that encompass the University of Toronto and then its surrounding areas.
+I then refined the neighbourhoods to include only those that encompass the University of Toronto and then its surrounding areas.
 The city of Toronto recognizes 158 unique neighbourhoods and of the 158, 13 have been included in the cleaned data.
@@
 Over the past decade, University of Toronto's encompassing neighbourhoods have exhibited unique trends in assault rates.
-"Kensington-Chinatown" shows an erratic pattern, peaking around 2018 followed by a sharp decline in 2019 and a subsequent rise again.
-"The Annex" demonstrates a generally consistent, however, slight upward trend.
+"Kensington-Chinatown" shows an erratic pattern, first peaking around 2018, followed by a sharp decline in 2019, and a subsequent rise again.
+"Kensington-Chinatown" appears to have experienced its highest rate of assault over the decade this past year, 2023.
+Overall, "Kensington-Chinatown" rates remain relatively high over the decade.
+The "Annex" demonstrates a generally consistent trend, however, in 2020, begins on an upward journey.
 Both "Bay-Cloverhill" and "University" show quite inconsistent patterns with several fluctuations over the ten year time span.
-What is most notable is Universities steady increase from 2020 on.
-The Annex also shows an increase from 2020 on, though not as steep as University.
+What is most notable is the steady increase in "University" from 2020 on.
+"Annex" also shows an increase from 2020 on, though not as steep as "University".
 @fig-assaults_per_year_uoft shows the assault rates per year for each University of Toronto neighbourhood.
@@
   labs(
-    title = "   Assault Rates in University of Toronto's 
-Surrounding Neighbourhoods (2014-2023)",
+    title = "Assault Rates in University of Toronto 
+    Neighbourhoods (2014-2023)",
     x = "Year",
@@
 Assault rates in neighbourhoods surrounding the University of Toronto reveals some distinct patterns and fluctuations.
-Several neighbourhoods exhibit fairly stable trends with minor year-to-year changes including "Yonge-St.Clair", "Wychwood", "West Queen West, "Palmerston-Little Italy", "Case Loma" and "Trinity-Bellwoods".
+Several neighbourhoods exhibit fairly stable trends with minor year-to-year changes including "Yonge-St.Clair", "Wychwood", "West Queen West", "Palmerston-Little Italy", "Casa Loma" and "Trinity-Bellwoods".
 "Dovercourt Village" shows a slight upward trend over the decade.
@@
 Clair", "Casa Loma", "Palmerston-Little Italy", and "Wychwood".
-For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to @fig-assaults_per_year_surounding_neighbourhoods.
+For a full picture of assault rates from 2014 to 2023 across various neighbourhoods surrounding the University of Toronto, please refer to @fig-assaults_per_year_surrounding_neighbourhoods.
 ``````{r fig-assaults_per_year_surounding_neighbourhoods, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Assault Rates in Surrounding Neighbourhoods"}
@@
 #| echo: false
-#| label: fig-assaults_per_year_surounding_neighbourhoods
+#| label: fig-assaults_per_year_surrounding_neighbourhoods
 #| fig-cap: Assault Rates in Surrounding Neighbourhoods (2014-2023)
@@
   labs(
-    title = "Assault Rates per year Across All Neighbourhoods From 2014-2023",
+    title = "Assault Rates per Year Across All Neighbourhoods (2014-2023)",
     x = "Year",
@@
 "University" is the neighbourhood predominantly occupied by the University of Toronto campus.
-The steady increase in assault rates in the "University" neighborhood since 2020, as shown in @fig-assaults_per_year_per_neighbourhoods, is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
+The steady increase in assault rates in the "University" neighbourhood since 2020, as shown in @fig-assaults_per_year_per_neighbourhoods, is particularly concerning when compared to the assault rates in the surrounding neighborhoods.
 Unlike the surrounding areas, which display relatively stable or minimal fluctuating rates over the same period, "University" exhibits a distinct and continuous rise.
-The next steepest rise in assault rates from 2020-2023 is "Annex", which is another recognized University of Toronto Neighbourhood.
-Additionally, assault rates in "Kensington-Chinatown" are most consistent and highest overall across the entirety of the decade then all surrounding neighnourhoods.
+The next steepest rise in assault rates from 2020-2023 is "Annex", which is another recognized University of Toronto neighbourhood.
+Additionally, assault rates in "Kensington-Chinatown" are most consistent and highest overall across the entirety of the decade than all surrounding neighbourhoods.
 These patterns within the University of Toronto neighbourhoods should raise concern and prompt conversation about assault rates on the University of Toronto campus.
 
-This divergence in trends between the "University" neighborhood and its neighboring areas suggests that the problem may be localized and thus require targeted interventions by the university.
+This divergence in trends between "University" and its neighbouring areas suggests that the problem may be localized and thus require targeted interventions by the university.
 These interventions could range from increasing campus security, bolstering support services for students, and implementing community engagement initiatives aimed at preventing assaults of all types.
@@
 # Conclusion
-This analysis of assault rates from 2014 to 2023 in and around the University of Toronto offers a unique view of campus safety in raletion to assault rates and its surrounding environment.
-The significant findings challenge the commonly held belief that university campuses, especially prestigious ones like the University of Toronto, are insulated and often safer then surrounding areas.
+This analysis of assault rates from 2014 to 2023 in and around the University of Toronto offers a unique view of campus safety in relation to assault rates and its surrounding environment.
+The significant findings challenge the commonly held belief that university campuses, especially prestigious ones like the University of Toronto, are insulated and often safer than surrounding areas.
 The data shows a notable increase in assault rates in the "University" and "Annex" neighborhoods, areas predominantly associated with the University of Toronto.
